---
title: "p8105_hw1_sef2183"
author: "Sarah Forrest"
date: "2022-09-16"
output: github_document
---
```{r load packages}
library(tidyverse)
```

# Problem 1
```{r load data}
data("penguins", package = "palmerpenguins")
summary(penguins)
```

## Dataset Description: 
* The 8 variables in the `penguins` dataset include: species, island, bill length (mm), bill depth (mm), flipper length (mm), body mass (g), sex, and year. The species variable is a factor variable with 3 levels: Adelie, Chinstrap, and Gentoo. The island variable is also a factor variable with 3 levels: Biscoe, Dream, and Torgersen The bill length and bill depth variables are numeric variables with 1 decimal place. The flipper length, body mass, and year variables are integer variables. Finally, the sex variable is a factor variable with 2 levels: male and female.

* The `penguins` dataset contains `r nrow(penguins)` observations (rows) and `r ncol(penguins)` variables (columns). 

* The mean flipper length in the `penguins` dataset is `r mean(penguins$flipper_length_mm, na.rm=TRUE)` mm.

## Scatterplot of Flipper Length (mm) vs. Bill Length (mm):
### Create Scatterplot
```{r create scatterplot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
```

### Save Scatterplot in p8105_hw1_sef2183 Project Directory
```{r save scatterplot}
ggsave("ScatterPlot.pdf", height = 4, width = 6)
```

# Problem 2

## Create the Dataframe "p2_df"
```{r create dataframe}
p2_df = 
  tibble(
    random = rnorm(n = 10),
    vec_logical = random > 0,
    vec_character = c("My", "name", "is", "Sarah", "and", "I", "live", "in", "Morningside", "Heights"),
    vec_factor = factor(c("yes", "no", "maybe", "yes", "maybe", "yes", "no", "no", "maybe", "no"))
  )
```

```{r view dataframe}
p2_df 
```
I created dataframe called "p2_df" with: 

* A random sample of size 10 from a standard normal distribution with the name "random"
* A logical vector indicating whether elements of the sample are greater than 0 with the name "vec_logical"
* A character vector of length 10 with the name "vec_character"
* A factor vector of length 10, with 3 different factor “levels” with the name "vec_factor"

## Check the Variable Means
### Mean of the random sample variable "random"
```{r check random mean}
mean(p2_df %>% pull(random))
```
### Mean of the logical vector "vec_logical"
```{r check logical mean}
mean(p2_df %>% pull(vec_logical))
```
### Mean of the character vector "vec_character"
```{r check character mean}
mean(p2_df %>% pull(vec_character))
```
### Mean of the factor vector "vec_factor"
```{r check factor mean}
mean(p2_df %>% pull(vec_factor))
```

When I try to take the mean of each variable in my "p2_df" dataframe:

* The mean of the random sample variable "random" is `r mean(p2_df$random)`
* The mean of the logical vector "vec_logical" is `r mean(p2_df$vec_logical)`
* The mean of the character vector "vec_character" returns `r mean(p2_df$vec_character)` because the argument is not numeric or logical
* The mean of the factor vector "vec_factor" returns `r mean(p2_df$vec_factor)` because the argument is not numeric or logical

## Convert Variable Types
```{r change variable types, echo = TRUE, results = 'hide', warning = FALSE}
as.numeric(p2_df$vec_logical)
as.numeric(p2_df$vec_character)
as.numeric(p2_df$vec_factor)
```

When I try to apply the `as.numeric` function to the logical, character, and factor variables:

* The logical vector "vec_logical" returns: `r as.numeric(p2_df$vec_logical)`, with 0s indicating values of `FALSE` and 1s indicating values of `TRUE`.
* The character vector "vec_character" returns: `r as.numeric(p2_df$vec_character)` because R cannot convert character variables to numeric, so it introduces NAs by coersion.
* The factor vector "vec_factor" returns: `r as.numeric(p2_df$vec_factor)`, with 1s indicating values of "maybe", 2s indicating values of "no", and 3s indicating values of "yes".

This helps explain why when I take the mean of the logical vector "vec_logical" R returns a value, while when I try to take the mean of the character vector "vec_character" R does not, and instead returns `r mean(p2_df$vec_character)`. 